Distinguishing Spin-Aligned and Isotropic Black Hole Populations With Gravitational Waves

### 2017-05-18 paper/AlignedIsotropicGW.tex
First pass fixing reference styles.  Fixes #21.

\usepackage{acronym}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{hyperref}
\usepackage{graphicx}

@@

\begin{abstract}
  The first direct detections of \acp{GW}
  \cite{2016PhRvL.116f1102A,2016PhRvL.116x1103A,O1-BBH} from merging
  \acp{BBH} open a unique window into the \ac{BBH} formation
  environment.  One promising signature of the formation environment
  is the angular distribution of the \ac{BH} spins; systems formed
  through dynamical interactions among already-compact objects are
  expected to have isotropic spin orientations
  \cite{SigurdssonHernquist:1993,1993Natur.364..421K,PZMcMillan:2000,Rodriguez:2015,Stone:2016,2016ApJ...832L...2R}
  whereas binaries formed from pairs of stars born together are more
  likely to have spins preferentially aligned with the binary orbit as
  a consequence of their joint evolution toward a \ac{BBH} system
  \cite{TutukovYungelson:1973,TutukovYungelson:1993,Lipunov:1997,2016Natur.534..512B,Stevenson:2017,MandeldeMink:2016,Marchant:2016}.
  We consider existing \ac{GW} measurements of $\chieff$, the
  best-measured combination of spin parameters
  \cite{2016PhRvL.116x1102A,O1-BBH}, in the three likely binary black
  hole detections GW150914, LVT151012, and GW151226.  If binary black
  hole spin magnitudes extend to high values, as is suggested by
  observations of black hole X-ray binaries
  \cite{2015PhR...548....1M}, we show that the data already exhibit a
  $\OOneSigmaIsoAligned\sigma$ ($\OOneOddsIsoAligned$ odds
  ratio\footnote{An odds ratio of $r$ with $r \ll 1$ is equivalent to
    $x \sigma$ with $x = \Phi^{-1}\left( 1 - r/2 \right)$, where
    $\Phi$ is the unit normal CDF.})  preference for an isotropic
  angular distribution.  By considering the effect of 10 additional
  detections \cite{2016ApJ...833L...1A} drawn from the various models
  in the suite we show that if all observations come from a single
  population such an augmented data set would enable at least a
  $\OTwoSigmaIsoAlignedMin\sigma$ ($\OTwoOddsIsoAlignedMin$ odds
  ratio) distinction between the isotropic and aligned models for the
  assumed spin magnitude distributions, and in most cases better than
  $5\sigma$ ($2.9 \times 10^{-7}$ odds ratio).  The existing
  preference for either an isotropic spin distribution or low spin
  magnitudes for the observed systems will be confirmed (or
  overturned) confidently in the near future by subsequent
  observations.
\end{abstract}

@@

Following the detection of a merging \ac{BBH} system, \ac{PE} tools
\cite{2015PhRvD..91d2003V} compare model gravitational waveforms
\cite{2014PhRvD..89h4006P,2014PhRvD..89f1502T,2014PhRvL.113o1101H}
against the observed data to obtain a posterior distribution on the
parameters that describe the compact binary source.

The spin parameter with the largest effect on waveforms, and a
correspondingly tight constraint from the data, is a mass-weighted,
aligned, ``effective spin,''
\begin{equation}
  \chieff = \frac{c}{GM} \left( \frac{\vec{S}_1}{m_1} + \frac{\vec{S}_2}{m_2}
  \right) \cdot \frac{\vec{L}}{\left| \vec{L} \right|} \equiv \frac{1}{M} \left( m_1 \chi_1 + m_2 \chi_2 \right),
\end{equation}
where $m_{1,2}$ are the gravitational masses of the more-massive (1)
and less-massive (2) components, $M = m_1 + m_2$ is the total mass,
$\vec{S}_{1,2}$ are the spin angular momentum vectors of the black
holes in the binary, $\vec{L}$ is the orbital angular momentum vector,
assumed to point in the $\hat{z}$ direction, and
$0 \leq \chi_{1,2} \leq 1$ are the corresponding dimensionless
projections of the individual \ac{BH} spins
\cite{2016PhRvL.116x1102A}.

Figure \ref{fig:O1-posteriors} shows an approximation to the posterior
inferred on $\chieff$ for the three likely \ac{GW} detections
GW150914, GW151226 and LVT151012 from Advanced LIGO's first observing
run (O1) \cite{O1-BBH}.  Because samples drawn from the posterior on
$\chieff$ are not publicly released at this time, we have approximated
the posterior as a Gaussian distribution with the same mean and 90\%
credible interval as quoted in \cite{O1-BBH}.  None of the $\chieff$
posteriors are consistent with two black holes with large aligned
spins, $\chi_{1,2} \gtrsim 0.5$; this contrasts with the large spins
inferred for the majority of black holes in X-ray binaries with claimed spin measurements (see
Section \ref{sec:discussion}).  The analysis here is relatively
insensitive to the precise details of the posterior distributions;
other conclusions are more sensitive.  In particular, our Gaussian
approximation does permit $\chieff = 0$ for GW151226 while the true
posterior rules this out at high confidence
\cite{2016PhRvL.116x1103A,O1-BBH}.

\begin{figure}
  \plotone{../plots/chi-eff-mock-posteriors}
  \caption{\label{fig:O1-posteriors} Approximate posteriors on
    $\chieff$ from the Advanced LIGO O1 run observations in
    \cite{O1-BBH}.  We approximate the posteriors reported in
    \cite{O1-BBH} using Gaussians with the same median and 90\%
    credible interval as reported in \cite{O1-BBH}.  It is notable
    that none of the $\chieff$ posteriors support high \ac{BH}
    spin magnitudes with aligned spins, suggested by observations of stellar-mass black holes in  X-ray binaries (see
    \cite{2015PhR...548....1M} for a summary of such measurements).}
\end{figure}

@@

Each of our models for the merging \ac{BBH} spin population assumes
that the merging black holes are of equal mass (this is marginally
consistent with the three observations \cite{O1-BBH}, and the
$\chieff$ distribution is not particularly sensitive to the mass ratio
between the merging objects---see Section \ref{sec:mass-ratio}).  We
assume that the population spin distribution factorises into a
distribution for the spin magnitude $a$ and a distribution for the
spin angles.  Finally, we assume that the distribution of spins is
common to each component in a merging binary.  Choosing one of three
magnitude distributions
\begin{equation}
  \label{eq:magnitude-dists}
  p(a) = \begin{cases}
    2\left(1-a \right) & \textnormal{``low''} \\
    1 & \textnormal{``flat''} \\
    2 a & \textnormal{``high''}
  \end{cases},
\end{equation}
and pairing with either an isotropic angular distribution or a
distribution that generates perfect alignment with the positive $z$
axis yields six different models for the $\chieff$ distribution.
These models are shown in Figure \ref{fig:chieff-distribution-models}.

@@

The distributions in Eq.\ \eqref{eq:magnitude-dists} are not meant to
represent any particular physical model, but rather to capture our
uncertainty about the spin magnitude distribution, discussed in detail
in Section \ref{sec:discussion}; neither observations nor population
synthesis codes can at this point authoritatively suggest \emph{any}
particular spin distribution \cite{2015PhR...548....1M}.  Our models,
however, allow us to see how sensitive the $\chieff$ distribution is
to spin alignment given uncertainties about the spin magnitudes.

We fit hierarchical models of the three LIGO O1 observations using
these six different, zero-parameter population distributions (see
Section \ref{sec:hierarchical}).  We also fit three mixture models for
the population, where the spin magnitude distribution is fixed but the
angular distribution is a weighted sum of the isotropic and aligned
distribution.  The evidence, or marginal likelihood, for each of the
models is shown in Figure \ref{fig:O1-odds}.  For all three magnitude
distributions, the mixture models' posterior on the mixing fraction
peaks at 100\% isotropic, which explains why the zero-parameter,
pure-isotropic models are preferred over the single-parameter mixture
models for every magnitude distribution with this data set.  Not
surprisingly, given the small $\chieff$ values in the three detected
systems, the most-favoured model among those with an isotropic angular
distribution has the ``low'' magnitude distribution; the most favoured
model among those with an aligned distribution also has the ``low''
magnitude distribution.  The odds ratio between the ``low'' aligned
and ``low'' isotropic models is $\OOneOddsIsoAligned$, or
$\OOneSigmaIsoAligned\sigma$; thus the data favour isotropic spins
among our suite of models.  While the data favour spin amplitude
distributions with small spin magnitudes, note that a model with all
\ac{BBH} systems having zero spin is ruled out by the GW151226
measurements, which bound at least one black hole to have spin
magnitude $\geq 0.2$ at 99\% credibility \cite{2016PhRvL.116x1103A}.

@@

Estimates of the rate of \ac{BBH} coalescences give a reasonable
chance of 10 additional \ac{BBH} detections in the next two years
\cite{O1-BBH,2016ApJ...833L...1A,2016ApJS..227...14A}.  Assuming 10
additional detections drawn from each of our six zero-parameter models
for the spin distribution in addition to the three existing detections
from O1, with observational uncertainties drawn randomly from the
three Gaussian widths used to approximate the $\chieff$ posteriors in
Figure \ref{fig:O1-posteriors}\footnote{The measurement uncertainty in
  $\chieff$ depends on the other parameters of the merging \ac{BBH}
  system, particularly on the mass ratio.  Our assumption about future
  observational uncertainties is appropriate if the parameters of the
  three detected events are representative of the parameters of future
  detections. See Section \ref{sec:chi-eff-precision}
for further discussion.}, we find the odds ratios shown in Figure
\ref{fig:O2-predictions}.  We find that most scenarios with an
additional 10 detections allow the simulated angular distribution to
be inferred with greater than $5\sigma$ ($2.9 \times 10^{-7}$ odds)
credibility; and in the most pessimistic case the distinction is
typically $\OTwoSigmaIsoAlignedMin\sigma$ ($\OTwoOddsIsoAlignedMin$
odds ratio).  While such future detections should permit a confident
distinction between \emph{angular} distributions, we would remain much
less certain about the \emph{magnitude} distribution among the three
options considered here until we have a larger number of observations.
In Figure \ref{fig:O2-predictions}, the odds ratio between different
magnitude distributions with the same angular distribution is much
closer to unity than the odds ratio between angular distributions.

@@

Most of our resolving power for the spin angular distribution is a
result of the fact that our ``aligned'' models cannot produce
$\chieff < 0$ (see Figure \ref{fig:chieff-distribution-models}).  If
spins are intrinsically very small, with $a \lesssim 0.3$, then it is
no longer possible to resolve the negative effective spin with a small
number of observations.  As noted below, however, spins observed in
X-ray binaries are typically large.  Additionally, models which do not
permit \emph{some} spins with $\chieff \gtrsim 0.1$ are ruled out by
the GW151226 observations \cite{2016PhRvL.116x1103A}.

In order to perform our analysis we need to select models for the
distribution of the spin magnitudes of stellar-mass black holes.
Observational data for such model selection is sparse.
\cite{2015PhR...548....1M} give current estimates of the spin
parameters for stellar-mass black holes, obtained using disk
reflection and/or disk continuum methods.  Most of the systems studied
are low-mass X-ray binaries rather than the high-mass X-ray binaries
that are likely to be the progenitors of double black hole binaries.
In addition, there are substantial systematic errors that can
complicate either type of analysis (see \cite{2015PhR...548....1M} for
a discussion), and selection effects could yield a biased
distribution. Nonetheless, if we take the reported spin magnitudes as
representative then we find that there is a preference for high spins;
for example, 14 of the 19 systems with reported spins have
dimensionless spin parameters in excess of 0.5.  It is usually argued
that the masses and spin parameters of stellar-mass black holes are
unlikely to be altered significantly by accretion (low-mass donors may
not have enough mass and high-mass donors have a very short phase in
which they transfer mass; a variant of this long-standing argument was
presented by \cite{1999MNRAS.305..654K}), but see
\cite{2003MNRAS.341..385P,2015ApJ...800...17F}.  Thus the current spin
parameters probably are close to their values upon core collapse, at
least in high-mass X-ray binaries.  However, the specific processes
involved in the production of black hole binaries from isolated
binaries could alter the spin magnitude distribution of those holes
relative to the X-ray binary systems; for example, close tidal
interactions could spin up the core, or stripping of the envelope
could reduce the available angular momentum
\cite{2016MNRAS.462..844K,2017arXiv170200885Z,2017arXiv170203952H}.

The spin directions of binary black holes formed dynamically through
interactions in dense stellar environments
\cite{SigurdssonHernquist:1993,1993Natur.364..421K,PZMcMillan:2000,Rodriguez:2015,Stone:2016}
are expected to be isotropic given the absence of a preferred
direction \cite{2016ApJ...832L...2R} and the persistence of an
isotropic distribution through post-Newtonian evolution
\cite{2004PhRvD..70l4020S,2007ApJ...661L.147B}.

The spin directions in isolated binaries, whether evolving via the
classical channel through a common-envelope phase
\cite{TutukovYungelson:1973,TutukovYungelson:1993,Lipunov:1997,2016Natur.534..512B,Stevenson:2017}
or through chemically homogeneous evolution
\cite{MandeldeMink:2016,Marchant:2016} are usually expected to be
preferentially aligned.  Despite observed spin-orbit misalignments in
both massive stellar binaries
\cite{Albrecht:2009,2014ApJ...785...83A} and BH X-ray binaries
\cite{Orosz:2001,Martin:2008b,Martin:2008,MorningstarMiller:2014},
mass accretion and tidal interactions will tend to realign the binary.
On the other hand, a supernova natal kick (if any) can change the
orbital plane and misalign the binary
\cite{2000ApJ...541..319K,2013PhRvD..87j4028G}; the supernova can
also tilt the spin angle, as in the double pulsar J0737-3039
\cite{2011ApJ...742...81F}; and a variety of uncertain processes,
such as wind-driven mass loss and post-collapse fallback, can couple
the spin magnitude and direction distributions, contrary to our
simplified assumptions.  A small misalignment at wide separation can
also evolve to a more significant misalignment in component spins as
the binary spirals in through \ac{GW} emission \cite{2015PhRvD..92f4016G},
but $\chieff$ is approximately conserved through this evolution.

\cite{2017CQGra..34cLT01V} also studied the possibility of
distinguishing aligned and isotropic angular distributions of \ac{BBH}
spins, but concluded that several hundred sources would be required to
adequately separate models which included both parallel and
anti-parallel spins for an ``aligned'' population.  In contrast, and
in agreement with the present study, \cite{Stevenson:2017spin} found
that only $\sim 5$ observations of rapidly spinning black holes would
be necessary to distinguish isotropic and aligned spin distributions,
though tens of detections would be required for more nuanced admixture
models.  Meanwhile, recent studies by
\cite{2017arXiv170306869F} and \cite{2017arXiv170306223G} focused on inference
on spin magnitudes.

@@

Following \cite{Stevenson:2017spin}, we fit a mixture model (labelled
model 'M' in Figure~\ref{fig:O1-odds}) where a fraction $f_i$ of BBHs
have spins drawn from an isotropic distribution, whilst a fraction
$1 - f_i$ have their spins aligned with the orbital angular
momentum. We assume a flat prior on the fraction $f_i$. To test the
robustness of our result, we vary the distribution we assume for BH
spin magnitude distributions as with the aligned and isotropic
models. We use the ``flat'', ``high'' and ``low'' distributions
(Equation~\ref{eq:magnitude-dists}), assuming all BHs have their spin
magnitude drawn from the same distribution for both the aligned and
isotropic populations. We calculate and plot the posterior on $f_i$
given by Equation~\ref{eq:hierarchical-posterior} ($f_i=\lambda$ in
the derivation) in Figure~\ref{fig:mixture_fraction_posterior}. We
find the mean fraction of BBHs coming from an isotropic distribution
is 0.63, 0.71 and 0.78 assuming the ``low'', ``flat'' and ``high''
distributions for spin magnitudes respectively, compared to the prior
mean of 0.5. The lower 90\% limits are 0.26, 0.39 and 0.52
respectively, compared to the prior of 0.1. In all cases, the
posterior peaks at $f_i = 1$. Thus, for these spin magnitude
distributions we find that the current O1 LIGO observations constrain
the majority of BBHs to have their spins drawn from an isotropic
distribution. The evidence ratios of these models to the isotropic
distribution with ``flat'' spin magnitudes are 0.85, 0.39 and 0.19 for
the ``low'', ``flat'' and ``high'' spin magnitude models. Thus we
cannot rule out a mixture with the current data.

@@

LIGO measures $\chieff$ better than any other spin parameter, but
still with significant uncertainty, so we need to properly incorporate
measurement uncertainty in our analysis; thus our analysis must be
\emph{hierarchical} \cite{2010ApJ...725.2166H,2010PhRvD..81h4029M}.
In a hierarchical analysis, we assume that each event has a true, but
unknown, value of the effective spin, drawn from the population
distribution, which may have some parameters $\lambda$; then the
system is observed, represented by the likelihood function, which
results in a distribution for the true effective spin (and all other
parameters describing the system) consistent with the data.
Combining, the joint posterior on each system's $\chieff^i$ parameters
and the population parameters $\lambda$ implied by a set of
observations each with data $d^i$, is
\begin{equation}
  p\left( \left\{ \chieff^i \right\}, \lambda \mid \left\{ d^i \right\} \right) \propto \left[ \prod_{i=1}^{N_\mathrm{obs}} p\left(d^i \mid \chieff^i \right) p\left( \chieff^i \mid \lambda \right) \right] p\left(\lambda\right).
\end{equation}

@@

\section{Approximations in the Gravitational Waveform}
The model waveforms used to infer the $\chieff$ of the three LIGO
events incorporate approximations to the true behaviour of the merging
systems that are expected to break down for sufficiently high
mis-aligned spins.  The effect of these approximations on inference on
the parameters describing GW150914 has been investigated in detail
\cite{2016arXiv161107531T}.  For this source, statistical uncertainties dominate
over any waveform systematics.  Detailed comparisons with numerical
relativity computations using no approximations to the dynamics
\cite{2016PhRvD..94f4035A} also suggest that statistical
uncertainties dominate the systematics for this system.
\cite{2016arXiv161107531T} suggests that systematics may dominate for
signals with this large SNR ($\simeq 23$) when the source is edge-on
or has high spins.  The other two events discussed in this paper are
at much lower SNR, with correspondingly larger statistical
uncertainties, and are probably similarly oriented and with similarly
small spins, so we do not expect systematic uncertainties to dominate.

>>>>

\usepackage{acronym}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{color}
\usepackage{hyperref}
\usepackage{graphicx}

@@

\begin{abstract}
  The first direct detections of
  \acp{GW}\cite{2016PhRvL.116f1102A,2016PhRvL.116x1103A,O1-BBH} from
  merging \acp{BBH} open a unique window into the \ac{BBH} formation
  environment.  One promising signature of the formation environment
  is the angular distribution of the \ac{BH} spins; systems formed
  through dynamical interactions among already-compact objects are
  expected to have isotropic spin
  orientations\cite{SigurdssonHernquist:1993,1993Natur.364..421K,PZMcMillan:2000,Rodriguez:2015,Stone:2016,2016ApJ...832L...2R}
  whereas binaries formed from pairs of stars born together are more
  likely to have spins preferentially aligned with the binary orbit as
  a consequence of their joint evolution toward a \ac{BBH}
  system\cite{TutukovYungelson:1973,TutukovYungelson:1993,Lipunov:1997,2016Natur.534..512B,Stevenson:2017,MandeldeMink:2016,Marchant:2016}.
  We consider existing \ac{GW} measurements of $\chieff$, the
  best-measured combination of spin
  parameters\cite{2016PhRvL.116x1102A,O1-BBH}, in the three likely
  binary black hole detections GW150914, LVT151012, and GW151226.  If
  binary black hole spin magnitudes extend to high values, as is
  suggested by observations of black hole X-ray
  binaries\cite{2015PhR...548....1M}, we show that the data already
  exhibit a $\OOneSigmaIsoAligned\sigma$ ($\OOneOddsIsoAligned$ odds
  ratio\footnote{An odds ratio of $r$ with $r \ll 1$ is equivalent to
    $x \sigma$ with $x = \Phi^{-1}\left( 1 - r/2 \right)$, where
    $\Phi$ is the unit normal CDF.})  preference for an isotropic
  angular distribution.  By considering the effect of 10 additional
  detections\cite{2016ApJ...833L...1A} drawn from the various models
  in the suite we show that if all observations come from a single
  population such an augmented data set would enable at least a
  $\OTwoSigmaIsoAlignedMin\sigma$ ($\OTwoOddsIsoAlignedMin$ odds
  ratio) distinction between the isotropic and aligned models for the
  assumed spin magnitude distributions, and in most cases better than
  $5\sigma$ ($2.9 \times 10^{-7}$ odds ratio).  The existing
  preference for either an isotropic spin distribution or low spin
  magnitudes for the observed systems will be confirmed (or
  overturned) confidently in the near future by subsequent
  observations.
\end{abstract}

@@

Following the detection of a merging \ac{BBH} system, \ac{PE}
tools\cite{2015PhRvD..91d2003V} compare model gravitational
waveforms\cite{2014PhRvD..89h4006P,2014PhRvD..89f1502T,2014PhRvL.113o1101H}
against the observed data to obtain a posterior distribution on the
parameters that describe the compact binary source.

The spin parameter with the largest effect on waveforms, and a
correspondingly tight constraint from the data, is a mass-weighted,
aligned, ``effective spin,''
\begin{equation}
  \chieff = \frac{c}{GM} \left( \frac{\vec{S}_1}{m_1} + \frac{\vec{S}_2}{m_2}
  \right) \cdot \frac{\vec{L}}{\left| \vec{L} \right|} \equiv \frac{1}{M} \left( m_1 \chi_1 + m_2 \chi_2 \right),
\end{equation}
where $m_{1,2}$ are the gravitational masses of the more-massive (1)
and less-massive (2) components, $M = m_1 + m_2$ is the total mass,
$\vec{S}_{1,2}$ are the spin angular momentum vectors of the black
holes in the binary, $\vec{L}$ is the orbital angular momentum vector,
assumed to point in the $\hat{z}$ direction, and
$0 \leq \chi_{1,2} \leq 1$ are the corresponding dimensionless
projections of the individual \ac{BH} spins\cite{2016PhRvL.116x1102A}.

Figure \ref{fig:O1-posteriors} shows an approximation to the posterior
inferred on $\chieff$ for the three likely \ac{GW} detections
GW150914, GW151226 and LVT151012 from Advanced LIGO's first observing
run (O1)\cite{O1-BBH}.  Because samples drawn from the posterior on
$\chieff$ are not publicly released at this time, we have approximated
the posterior as a Gaussian distribution with the same mean and 90\%
credible interval.  None of the $\chieff$ posteriors are consistent
with two black holes with large aligned spins,
$\chi_{1,2} \gtrsim 0.5$; this contrasts with the large spins inferred
for the majority of black holes in X-ray binaries with claimed spin
measurements (see Section \ref{sec:discussion}).  The analysis here is
relatively insensitive to the precise details of the posterior
distributions; other conclusions are more sensitive.  In particular,
our Gaussian approximation does permit $\chieff = 0$ for GW151226
while the true posterior rules this out at high
confidence\cite{2016PhRvL.116x1103A,O1-BBH}.

\begin{figure}
  \plotone{../plots/chi-eff-mock-posteriors}
  \caption{\label{fig:O1-posteriors} Approximate posteriors on
    $\chieff$ from the Advanced LIGO O1 run observations\cite{O1-BBH}.
    We approximate the posteriors reported using Gaussians with the
    same median and 90\% credible interval.  It is notable that none
    of the $\chieff$ posteriors support high \ac{BH} spin magnitudes
    with aligned spins, suggested by observations of stellar-mass
    black holes in X-ray binaries\cite{2015PhR...548....1M}.}
\end{figure}

@@

Each of our models for the merging \ac{BBH} spin population assumes
that the merging black holes are of equal mass (this is marginally
consistent with the three observations\cite{O1-BBH}, and the
$\chieff$ distribution is not particularly sensitive to the mass ratio
between the merging objects---see Section \ref{sec:mass-ratio}).  We
assume that the population spin distribution factorises into a
distribution for the spin magnitude $a$ and a distribution for the
spin angles.  Finally, we assume that the distribution of spins is
common to each component in a merging binary.  Choosing one of three
magnitude distributions
\begin{equation}
  \label{eq:magnitude-dists}
  p(a) = \begin{cases}
    2\left(1-a \right) & \textnormal{``low''} \\
    1 & \textnormal{``flat''} \\
    2 a & \textnormal{``high''}
  \end{cases},
\end{equation}
and pairing with either an isotropic angular distribution or a
distribution that generates perfect alignment with the positive $z$
axis yields six different models for the $\chieff$ distribution.
These models are shown in Figure \ref{fig:chieff-distribution-models}.

@@

The distributions in Eq.\ \eqref{eq:magnitude-dists} are not meant to
represent any particular physical model, but rather to capture our
uncertainty about the spin magnitude distribution, discussed in detail
in Section \ref{sec:discussion}; neither observations nor population
synthesis codes can at this point authoritatively suggest \emph{any}
particular spin distribution\cite{2015PhR...548....1M}.  Our models,
however, allow us to see how sensitive the $\chieff$ distribution is
to spin alignment given uncertainties about the spin magnitudes.

We fit hierarchical models of the three LIGO O1 observations using
these six different, zero-parameter population distributions (see
Section \ref{sec:hierarchical}).  We also fit three mixture models for
the population, where the spin magnitude distribution is fixed but the
angular distribution is a weighted sum of the isotropic and aligned
distribution.  The evidence, or marginal likelihood, for each of the
models is shown in Figure \ref{fig:O1-odds}.  For all three magnitude
distributions, the mixture models' posterior on the mixing fraction
peaks at 100\% isotropic, which explains why the zero-parameter,
pure-isotropic models are preferred over the single-parameter mixture
models for every magnitude distribution with this data set.  Not
surprisingly, given the small $\chieff$ values in the three detected
systems, the most-favoured model among those with an isotropic angular
distribution has the ``low'' magnitude distribution; the most favoured
model among those with an aligned distribution also has the ``low''
magnitude distribution.  The odds ratio between the ``low'' aligned
and ``low'' isotropic models is $\OOneOddsIsoAligned$, or
$\OOneSigmaIsoAligned\sigma$; thus the data favour isotropic spins
among our suite of models.  While the data favour spin amplitude
distributions with small spin magnitudes, note that a model with all
\ac{BBH} systems having zero spin is ruled out by the GW151226
measurements, which bound at least one black hole to have spin
magnitude $\geq 0.2$ at 99\% credibility\cite{2016PhRvL.116x1103A}.

@@

Estimates of the rate of \ac{BBH} coalescences give a reasonable
chance of 10 additional \ac{BBH} detections in the next two
years\cite{O1-BBH,2016ApJ...833L...1A,2016ApJS..227...14A}.  Assuming
10 additional detections drawn from each of our six zero-parameter
models for the spin distribution in addition to the three existing
detections from O1, with observational uncertainties drawn randomly
from the three Gaussian widths used to approximate the $\chieff$
posteriors in Figure \ref{fig:O1-posteriors}\footnote{The measurement
  uncertainty in $\chieff$ depends on the other parameters of the
  merging \ac{BBH} system, particularly on the mass ratio.  Our
  assumption about future observational uncertainties is appropriate
  if the parameters of the three detected events are representative of
  the parameters of future detections. See Section
  \ref{sec:chi-eff-precision} for further discussion.}, we find the
odds ratios shown in Figure \ref{fig:O2-predictions}.  We find that
most scenarios with an additional 10 detections allow the simulated
angular distribution to be inferred with greater than $5\sigma$
($2.9 \times 10^{-7}$ odds) credibility; and in the most pessimistic
case the distinction is typically $\OTwoSigmaIsoAlignedMin\sigma$
($\OTwoOddsIsoAlignedMin$ odds ratio).  While such future detections
should permit a confident distinction between \emph{angular}
distributions, we would remain much less certain about the
\emph{magnitude} distribution among the three options considered here
until we have a larger number of observations.  In Figure
\ref{fig:O2-predictions}, the odds ratio between different magnitude
distributions with the same angular distribution is much closer to
unity than the odds ratio between angular distributions.

@@

Most of our resolving power for the spin angular distribution is a
result of the fact that our ``aligned'' models cannot produce
$\chieff < 0$ (see Figure \ref{fig:chieff-distribution-models}).  If
spins are intrinsically very small, with $a \lesssim 0.3$, then it is
no longer possible to resolve the negative effective spin with a small
number of observations.  As noted below, however, spins observed in
X-ray binaries are typically large.  Additionally, models which do not
permit \emph{some} spins with $\chieff \gtrsim 0.1$ are ruled out by
the GW151226 observations\cite{2016PhRvL.116x1103A}.

In order to perform our analysis we need to select models for the
distribution of the spin magnitudes of stellar-mass black holes.
Observational data for such model selection is
sparse\cite{2015PhR...548....1M}.  Most of the systems studied are
low-mass X-ray binaries rather than the high-mass X-ray binaries that
are likely to be the progenitors of double black hole binaries.  In
addition, there are substantial systematic errors that can complicate
either type of analysis\cite{2015PhR...548....1M} and selection
effects could yield a biased distribution. Nonetheless, if we take the
reported spin magnitudes as representative then we find that there is
a preference for high spins; for example, 14 of the 19 systems with
reported spins have dimensionless spin parameters in excess of 0.5.
It is usually argued that the masses and spin parameters of
stellar-mass black holes are unlikely to be altered significantly by
accretion (low-mass donors may not have enough mass and high-mass
donors have a very short phase in which they transfer
mass\cite{1999MNRAS.305..654K}), but this may not be true for all
systems\cite{2003MNRAS.341..385P,2015ApJ...800...17F}.  Thus the
current spin parameters probably are close to their values upon core
collapse, at least in high-mass X-ray binaries.  However, the specific
processes involved in the production of black hole binaries from
isolated binaries could alter the spin magnitude distribution of those
holes relative to the X-ray binary systems; for example, close tidal
interactions could spin up the core, or stripping of the envelope
could reduce the available angular
momentum\cite{2016MNRAS.462..844K,2017arXiv170200885Z,2017arXiv170203952H}.

The spin directions of binary black holes formed dynamically through
interactions in dense stellar
environments\cite{SigurdssonHernquist:1993,1993Natur.364..421K,PZMcMillan:2000,Rodriguez:2015,Stone:2016}
are expected to be isotropic given the absence of a preferred
direction\cite{2016ApJ...832L...2R} and the persistence of an
isotropic distribution through post-Newtonian
evolution\cite{2004PhRvD..70l4020S,2007ApJ...661L.147B}.

The spin directions in isolated binaries, whether evolving via the
classical channel through a common-envelope
phase\cite{TutukovYungelson:1973,TutukovYungelson:1993,Lipunov:1997,2016Natur.534..512B,Stevenson:2017}
or through chemically homogeneous
evolution\cite{MandeldeMink:2016,Marchant:2016} are usually expected
to be preferentially aligned.  Despite observed spin-orbit
misalignments in both massive stellar
binaries\cite{Albrecht:2009,2014ApJ...785...83A} and BH X-ray
binaries\cite{Orosz:2001,Martin:2008b,Martin:2008,MorningstarMiller:2014},
mass accretion and tidal interactions will tend to realign the binary.
On the other hand, a supernova natal kick (if any) can change the
orbital plane and misalign the
binary\cite{2000ApJ...541..319K,2013PhRvD..87j4028G}; the supernova
can also tilt the spin angle, as in the double pulsar
J0737-3039\cite{2011ApJ...742...81F}; and a variety of uncertain
processes, such as wind-driven mass loss and post-collapse fallback,
can couple the spin magnitude and direction distributions, contrary to
our simplified assumptions.  A small misalignment at wide separation
can also evolve to a more significant misalignment in component spins
as the binary spirals in through \ac{GW}
emission\cite{2015PhRvD..92f4016G}, but $\chieff$ is approximately
conserved through this evolution.

\fixme{\cite{2017CQGra..34cLT01V} also studied the possibility of
  distinguishing aligned and isotropic angular distributions of
  \ac{BBH} spins, but concluded that several hundred sources would be
  required to adequately separate models which included both parallel
  and anti-parallel spins for an ``aligned'' population.  In contrast,
  and in agreement with the present study,\cite{Stevenson:2017spin}
  found that only $\sim 5$ observations of rapidly spinning black
  holes would be necessary to distinguish isotropic and aligned spin
  distributions, though tens of detections would be required for more
  nuanced admixture models.  Meanwhile, recent studies
  by\cite{2017arXiv170306869F} and\cite{2017arXiv170306223G} focused
  on inference on spin magnitudes.}

@@

We also fit a mixture model\cite{Stevenson:2017spin} (labelled model
'M' in Figure~\ref{fig:O1-odds}) where a fraction $f_i$ of BBHs have
spins drawn from an isotropic distribution, whilst a fraction
$1 - f_i$ have their spins aligned with the orbital angular
momentum. We assume a flat prior on the fraction $f_i$. To test the
robustness of our result, we vary the distribution we assume for BH
spin magnitude distributions as with the aligned and isotropic
models. We use the ``flat'', ``high'' and ``low'' distributions
(Equation~\ref{eq:magnitude-dists}), assuming all BHs have their spin
magnitude drawn from the same distribution for both the aligned and
isotropic populations. We calculate and plot the posterior on $f_i$
given by Equation~\ref{eq:hierarchical-posterior} ($f_i=\lambda$ in
the derivation) in Figure~\ref{fig:mixture_fraction_posterior}. We
find the mean fraction of BBHs coming from an isotropic distribution
is 0.63, 0.71 and 0.78 assuming the ``low'', ``flat'' and ``high''
distributions for spin magnitudes respectively, compared to the prior
mean of 0.5. The lower 90\% limits are 0.26, 0.39 and 0.52
respectively, compared to the prior of 0.1. In all cases, the
posterior peaks at $f_i = 1$. Thus, for these spin magnitude
distributions we find that the current O1 LIGO observations constrain
the majority of BBHs to have their spins drawn from an isotropic
distribution. The evidence ratios of these models to the isotropic
distribution with ``flat'' spin magnitudes are 0.85, 0.39 and 0.19 for
the ``low'', ``flat'' and ``high'' spin magnitude models. Thus we
cannot rule out a mixture with the current data.

@@

LIGO measures $\chieff$ better than any other spin parameter, but
still with significant uncertainty, so we need to properly incorporate
measurement uncertainty in our analysis; thus our analysis must be
\emph{hierarchical}\cite{2010ApJ...725.2166H,2010PhRvD..81h4029M}.
In a hierarchical analysis, we assume that each event has a true, but
unknown, value of the effective spin, drawn from the population
distribution, which may have some parameters $\lambda$; then the
system is observed, represented by the likelihood function, which
results in a distribution for the true effective spin (and all other
parameters describing the system) consistent with the data.
Combining, the joint posterior on each system's $\chieff^i$ parameters
and the population parameters $\lambda$ implied by a set of
observations each with data $d^i$, is
\begin{equation}
  p\left( \left\{ \chieff^i \right\}, \lambda \mid \left\{ d^i \right\} \right) \propto \left[ \prod_{i=1}^{N_\mathrm{obs}} p\left(d^i \mid \chieff^i \right) p\left( \chieff^i \mid \lambda \right) \right] p\left(\lambda\right).
\end{equation}

@@

\section{Approximations in the Gravitational Waveform}
The model waveforms used to infer the $\chieff$ of the three LIGO
events incorporate approximations to the true behaviour of the merging
systems that are expected to break down for sufficiently high
mis-aligned spins.  The effect of these approximations on inference on
the parameters describing GW150914 has been investigated in
detail\cite{2016arXiv161107531T}.  For this source, statistical
uncertainties dominate over any waveform systematics.  Detailed
comparisons with numerical relativity computations using no
approximations to the dynamics\cite{2016PhRvD..94f4035A} also suggest
that statistical uncertainties dominate the systematics for this
system.  Systematics may dominate for signals with this large SNR
($\simeq 23$) when the source is edge-on or has high
spins\cite{2016arXiv161107531T}.  The other two events discussed in
this paper are at much lower SNR, with correspondingly larger
statistical uncertainties, and are probably similarly oriented and
with similarly small spins, so we do not expect systematic
uncertainties to dominate.
